Scoring epidemiological forecasts on transformed scales

### 2023-08-04 final submission after reformatting/final-manuscript.tex
Adapt formatting for final submission

\documentclass{article}
\usepackage[]{graphicx}
\usepackage[]{xcolor}
\usepackage{alltt}
\usepackage[left=2.3cm,right=2.8cm, top = 2.2cm, bottom = 3cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\PassOptionsToPackage{hyphens}{url}
\usepackage{url} 
\usepackage[disable]{todonotes}
\usepackage{multicol}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage[colorlinks=false]{hyperref} 
\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%
\urlstyle{same}
\usepackage{lineno}
\linenumbers
\bibliographystyle{apalike}
%\bibliographystyle{vancouver}

@@

\begin{document}

\title{Scoring epidemiological forecasts on transformed scales}
  \author{Nikos I. Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,  
  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$, 
  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$, 
  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}

\maketitle
\begin{abstract}

Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation. However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
\end{abstract}

\bigskip

{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}, 
{\footnotesize $^\dagger$ Contributed equally}

\section*{Author summary}
Scores like the Continuous Ranked Probability Score (CRPS) or the Weighted Interval Score (WIS) are commonly used to evaluate epidemiological forecasts and are a measure of absolute distance between forecast and observation. Due to the exponential nature of epidemic processes, evaluating the absolute distance between forecast and observation may not be ideal. We argue that transforming counts before applying the CRPS or WIS can yield more meaningful results. The natural logarithm is a particularly attractive transformation in epidemiological settings. Scores computed on log-transformed values can be interpreted as a probabilistic version of a relative error and reflect how well forecasters predict the time-varying epidemic growth rate. If the data-generating process has a quadratic mean-variance relationship, the logarithmic transformation also leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. 
We illustrate these properties using data from the European COVID-19 Forecast Hub and find that scoring transformed counts changes model rankings. Stronger emphasis is given to  situations in which forecasters missed the beginning of upward swings, while failing to predict a downturn following a peak is less severely penalised. We generally recommend including evaluations of transformed counts when assessing forecaster performance.

% ===========================================================
\section{Introduction}

Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}. Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic. Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles. They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.

Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. 
Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score. A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores. Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future. 
Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}. The log score is the predictive log density or probability mass evaluated at the observed value. It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}. Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts. It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.

The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
\end{equation}    
\end{linenomath*}
where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided. The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.

The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread. Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}. If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$. Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is

@@

For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena. We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation). In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$. For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple. For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
%
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
\end{equation}    
\end{linenomath*}
%
Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
\begin{linenomath*}
\begin{equation}
F_{\log}(x) = F(\exp(x)),
\end{equation}    
\end{linenomath*}
with $F$ the CDF on the original scale. Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}). 
We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic. We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}). The question of the right scale to evaluate forecasts on does therefore not arise for the log score.

The remainder of the article is structured as follows. In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}). To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}. Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Logarithmic transformation of forecasts and observations}
\label{sec:methods}

\subsection{Interpretation as a relative error}
\label{sec:methods:relative}

>>>>

% Template for PLoS
% Version 3.6 Aug 2022
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage[nopatch=eqnum]{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

%% Custom packages
\usepackage{booktabs}
% \usepackage[]{xcolor}
% \usepackage{alltt}
% \usepackage{amsmath}
% \usepackage{amssymb}
% \PassOptionsToPackage{hyphens}{url}
% \usepackage{url} 
% \usepackage[disable]{todonotes}
% \usepackage{multicol}
% \usepackage{rotating}

% \setlength{\parskip}{\baselineskip}%
% \setlength{\parindent}{0pt}%
% \urlstyle{same}
% \usepackage{lineno}
% \linenumbers

% \bibliographystyle{apalike}
%\bibliographystyle{vancouver}

@@

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Scoring epidemiological forecasts on transformed scales} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nikos I. Bosse\textsuperscript{1, 2, 3, *},
Sam Abbott\textsuperscript{1, 2},
Anne Cori\textsuperscript{4},
Edwin van Leeuwen\textsuperscript{1, 3, 5},
Johannes Bracher\textsuperscript{6, 7, \Yinyang},
Sebastian Funk\textsuperscript{1, 2, 3, \Yinyang}
\\
\bigskip
\textbf{1} Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom
\\
\textbf{2} Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom
\\
\textbf{3} NIHR Health Protection Research Unit in Modelling \& Health Economics
\\
\textbf{4} MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom
\\
\textbf{5} Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom
\\
\textbf{6} Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany
\\
\textbf{7} Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* nikos.bosse@lshtm.ac.uk

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation. However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Scores like the Continuous Ranked Probability Score (CRPS) or the Weighted Interval Score (WIS) are commonly used to evaluate epidemiological forecasts and are a measure of absolute distance between forecast and observation. Due to the exponential nature of epidemic processes, evaluating the absolute distance between forecast and observation may not be ideal. We argue that transforming counts before applying the CRPS or WIS can yield more meaningful results. The natural logarithm is a particularly attractive transformation in epidemiological settings. Scores computed on log-transformed values can be interpreted as a probabilistic version of a relative error and reflect how well forecasters predict the time-varying epidemic growth rate. If the data-generating process has a quadratic mean-variance relationship, the logarithmic transformation also leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. 
We illustrate these properties using data from the European COVID-19 Forecast Hub and find that scoring transformed counts changes model rankings. Stronger emphasis is given to  situations in which forecasters missed the beginning of upward swings, while failing to predict a downturn following a peak is less severely penalised. We generally recommend including evaluations of transformed counts when assessing forecaster performance.

\linenumbers

\section*{Introduction}

Probabilistic forecasts \cite{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \cite{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \cite{timmermannForecastingMethodsFinance2018} or meteorology \cite{gneitingWeatherForecastingEnsemble2005}. Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic. Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles. They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.

Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \cite{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. 
Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score. A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores. Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future. 
Common proper scoring rules are the logarithmic or log score \cite{goodRationalDecisions1952} and the continuous ranked probability score (CRPS,~\cite{gneitingStrictlyProperScoring2007}). The log score is the predictive log density or probability mass evaluated at the observed value. It is supported by the likelihood principle \cite{winklerScoringRulesEvaluation1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \cite{bracherEvaluatingEpidemicForecasts2021}. Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts. It is nonetheless occasionally used in epidemiology (see e.g., \cite{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score (WIS,~\cite{bracherEvaluatingEpidemicForecasts2021}) have become increasingly popular.

The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
\end{equation}    
\end{linenomath*}

where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \cite{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \cite{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \cite{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\cite{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided. The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.

The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread. Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \cite{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \cite{dushoffSpeedStrengthEpidemic2021}. If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$. Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is

@@

For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena. We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation). In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$. For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple. For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
%
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
\end{equation}    
\end{linenomath*}
%
Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
\begin{linenomath*}
\begin{equation}
F_{\log}(x) = F(\exp(x)),
\end{equation}    
\end{linenomath*}
with $F$ the CDF on the original scale. Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}). 
We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic. We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \cite{Lehmann1950} on corresponding properties of likelihood ratios and \cite{diksLikelihoodbasedScoringRules2011}). The question of the right scale to evaluate forecasts on does therefore not arise for the log score.

The remainder of the article is structured as follows. In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}). To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  (\cite{europeancovid-19forecasthubEuropeanCovid19Forecast2021,sherrattPredictivePerformanceMultimodel2022}, Section \ref{sec:HUB}). Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Logarithmic transformation of forecasts and observations}
\label{sec:methods}

\subsection*{Interpretation as a relative error}
\label{sec:methods:relative}
